---
title: "Many Local Maxima and Modeling Beetle Data Problems"
subtitle: "5361 Homework 4"
author: Qinxiao Shi ^[<qinxiao.shi@uconn.edu>]
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::pdf_document2
documentclass: article
papersize: letter
fontsize: 11pt
bibliography: Format.bib
biblio-style: asa
---

# Many Local Maxima
## Log-likelihood Function
The probability density function with parameter $\theta$ is:
$$f(x;\theta)= \frac{1-\cos(x-\theta)}{2\pi},\space\space 0 \le x \le 2\pi,\space\space \theta \in (-\pi , \pi)$$
\end{equation*}
The likelihood function of $f(x;\theta)$ is:
$$L(\theta)=\frac{\prod_{i=1}^{n}[1-\cos(X_i-\theta)]}{(2\pi)^n}$$
The log-likelihood function is:
$$\ell(\theta)=-n\log (2\pi)+\sum_{i=1}^{n}\log[1-\cos(X_i-\theta)]$$
```{r likelihood, echo=TRUE}
library("ggplot2")
set.seed(20180909)

x <- c(3.91, 4.85, 2.28, 4.06, 3.70, 4.04, 5.46, 3.53, 2.28, 1.96,
       2.53, 3.88, 2.22, 3.47, 4.82, 2.46, 2.99, 2.54, 0.52)

y <- function(theta){
  y <- 0
  for (i in 1:length(x)){
    y <- y-log(2*pi)+log(1-cos(x[i]-theta))
  }
  return(y)
}

ggplot(data.frame(theta=c(-pi,pi)), aes(x=theta))+
  stat_function(fun = function(theta) y(theta))+
  ggtitle(expression("Loglikelihood Funciton VS."~theta))+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(y="Value of Log-likelihood Funciton", x=expression(theta))
```

## Method-of-Moments Estimator
The expectation of $f(x;\theta)$ is:
\begin{equation*}
  \begin{split}
    \mathbb {E} (X \ |\theta)&=\int_{0}^{2\pi} xf(x;\theta)dx \\
    &=\frac{1}{2 \pi}(\int_{0}^{2\pi}xdx-\int_{0}^{2\pi}x\cos(x-\theta)dx) \\
    &=\frac{1}{2 \pi}[2\pi^2-(x\sin(x-\theta)\big|_{0}^{2\pi}-\int_{0}^{2\pi}\sin(x-\theta)dx)]\\
    &=\frac{1}{2 \pi}[2\pi^2-(x\sin(x-\theta)+\cos(x-\theta))\bigg|_{0}^{2\pi}]\\
    &=\frac{1}{2 \pi}\{ 2\pi^2-[x(\sin(x)\cos(\theta)-\cos(x)\sin(\theta))+\cos(x)\cos(\theta)+\sin(x)\sin(\theta)\bigg|_{0}^{2\pi})]\}\\
    &=\frac{1}{2 \pi}[2\pi^2+2\pi\sin{\theta}]\\
    &=\pi+\sin{\theta}
  \end{split}
\end{equation*}

So when we do estimation:
$\bar{X_{n}}=\pi+\sin{\hat\theta}$,  $\tilde{\theta}_{n}=\arcsin(\bar{X_{n}}-\pi)=arcsin(3.2368-\pi)=0.0954$

## Newton–Raphson Method For MLE
The gradient is:
$$\ell^\prime(\theta)=\sum_{i=1}^{n}\frac{-\sin(X_i-\theta)}{1-\cos(X_i-\theta)}$$
The hessian is:
\begin{equation*}
  \begin{split}
  \ell^{\prime\prime}(\theta)&=\sum_{i=1}^{n}\frac{-\cos(X_i-\theta)[1-\cos(X_i-\theta)]+[\sin(X_i-\theta)]^2}{[1-\cos(X_i-\theta)]^2}\\
  &=\sum_{i=1}^{n}\frac{-\cos(X_i-\theta)+1}{[1-\cos(X_i-\theta)]^2}\\
  &=\sum_{i=1}^{n}\frac{1}{1-\cos(X_i-\theta)}
  \end{split}
\end{equation*}

```{r MLE, echo=TRUE, warning=FALSE}
library("pracma")
library("pander")
library("gridExtra")
library("grid")
library("knitr")
library("kableExtra")

gradient <- function(theta){
    gradient <- sum(-sin(x-theta)/(1-cos(x-theta)))
  return(gradient)
}

hessian <- function(theta){
    hessian <- sum(1/(cos(x-theta)-1))
  return(hessian)
}

theta0 <- asin(mean(x)-pi)

newton1 <- newtonRaphson(fun=function(theta) gradient(theta), x0=theta0,
                         dfun=function(theta) hessian(theta))
  root1 <- newton1$root
  
table1 <- data.frame(Theta=theta0, Root=root1)

kable(table1, booktabs = TRUE, align = 'c', row.names = 1)


```
So the $MLE$ for $\theta$ using the Newton-Raphson method with initial value $\theta_0=\tilde\theta_n$ is $0.0031$.

## Find MLE at $\theta_0=\pm2.7$
```{r IA, echo=TRUE, warning=FALSE}
theta2 <- c(-2.7, 2.7)
newton2 <- vector("list", length = length(theta2))
root2 <- array(NA, dim=length(theta2))
for (i in 1:length(theta2)){
  newton2[[i]] <- newtonRaphson(fun=function(theta) gradient(theta), 
                              dfun=function(theta) hessian(theta), x0=theta2[i])
  root2[i] <- newton2[[i]]$root
}

table2 <- data.frame(Theta=theta2, Root=root2)
kable(table2, booktabs = TRUE, align = 'c', row.names = 1)
```
So the $MLE$ for $\theta$ at $\theta_0=-2.7$ is $-2.6689$, and the $MLE$ at $\theta_0=2.7$ is $2.8484$.

## 200 Values Optimization Between $-\pi$ and $\pi$
```{r optimization, echo=TRUE, warning=FALSE}
dff <- 2*pi/199
theta3 <- array(-pi, dim = 200)
for (i in 2:200)
  theta3[i] <- theta3[as.numeric(i-1)]+dff

newton3 <- vector("list", length = length(theta3))
root3 <- array(NA, dim=length(theta3))

for (i in 1:length(theta3)){
  newton3[[i]] <- newtonRaphson(fun=function(theta) gradient(theta), 
                              dfun=function(theta) hessian(theta), x0=theta3[i])
  root3[i] <- newton3[[i]]$root
}
  table3 <- data.frame(Theta=theta3, Root=root3)
  
temp <- as.data.frame(table(table3$Root))
mult.group <- vector("list", length(temp$Freq))
pos <- 0

for (i in 1:length(temp$Freq)) {
  for (j in 1:temp$Freq[i]){
  mult.group [[i]][j] <- table3$Theta[pos + j]
  }
  pos <- pos + temp$Freq[i]
}
```

So the group of initial value, whose local maximam are the same, should be:
```{r result, echo=TRUE, warning=FALSE}
print(mult.group)
  
  ggplot(table3, aes(x = theta3, y = root3)) + 
    geom_point()+
    ggtitle("Local Maximam VS."~theta)+
    theme(plot.title = element_text(hjust = 0.5))+
    labs(y="Local Maximam Value", x=expression(theta))
```

# Modeling Beetle Data
## Population Growth Model - Gauss-Newton Approach
```{r model, echo=TRUE, warning=FALSE}
beetles <- data.frame(
    days    = c(0,  8,  28,  41,  63,  69,   97, 117,  135,  154),
    beetles = c(2, 47, 192, 256, 768, 896, 1120, 896, 1184, 1024))

bt <- list(
t <- beetles$days,
y <- beetles$beetles
)

fo <- y~2*K/(2+(K-2)*exp(-r*t))
nls(fo, data=bt, start = list(K=1000, r=1))
```
The fitted model should be:
$$f(t)=\frac{2098.8136}{2+1047.4068\exp(-0.1183t)}$$
where the minimized sum of squared error should be $73420$

```{r contour, echo=TRUE, warning=FALSE}
K <- seq(500, 1500, by=5)
r <- seq(0, 1, by=0.005)


z <- matrix(NA, nrow = 201, ncol = 201)
for (i in 1:201){
  for (j in 1:201){
    z[i, j] <- sum((y-2*K[i]/(2+(K[i]-2)*exp(-r[j]*t)))^2)
  }
}
contour(K, r, z, xlab = "r", ylab = "K")
```
